A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2019-02-19 reviewer_response.tex

Associate Editor

Fluid Phase Equlibria
\vspace{1cm}

Dear Prof. Potoff:

Thank you for your message forwarding the reviewers’ comments on our manuscript ``A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquid \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}''.
We have carefully addressed the comments and suggestions, which required us to make changes in the text.
The most substantial revision took place in Sec.~5.3, for which we carried out additional simulations and included new figures.
Besides, we have included the full names of the ionic liquids in the abstract.
To facilitate your evaluation, we attached an annotated version in which all removed text has been crossed out in red and all new content is colored in blue.

@@

{\color{blue}{1) The authors chose not to use an iterative scheme to correct the pressure of their simulations. They should justify their choice by including quantitative information from their simulations on how this correction would affect both the differences in the free energy but also the system's density at high concentrations.}}

\textbf{Answer}:
In order to address the reviewer's concern on this topic, we decided to carry out a more rigorous strategy to compute the equilibrium pressure.
Instead of performing an iterative scheme, we executed some simulations at different pressures in the vicinity of the experimental value.
This allowed us to employ a multistate analysis method (MBAR) and compute the solvation free energy along a continuous range of pressures via reweighting.
The resulting curves confirmed our previous assertion that such free energy depends only slightly on the pressure.
Besides, we could use these curves to find, for each case, the pressure at which the equilibrium equations are satisfied.
We believe that the new procedure, which is explained on page 4 of the annotated manuscript, improves our analysis of the equilibrium condition in a significant way.

{\color{blue}{2) Among the various existing models used in this work, some explicitly report the use of LJ long range corrections in the energy and pressure. The authors should clearly state if they have taken that into account in their simulations.}}

\textbf{Answer}: The reviewer is right. We applied standard tail corrections to the LJ interactions for all systems under investigation. This information was included on page 5 of the manuscript.

@@

{\color{blue}{3) Results on densities of the various IL models at 298 K (Table 1) differ from the ones reported in the initial references. For example,  an 1\% error is presented in this manuscript for the FF-3 model of Koller \textit{et al.}, while an excellent agreement with the experimental density is reported in the original paper of Koller and his co-workers. The authors should explain this discrepancy.}}

\textbf{Answer}: The differences in density between our results and those reported by Koller \textit{et al}. \cite{Koller_2012}, Batista \textit{et al}. \cite{Batista_2015}, Liu \textit{et al}. \cite{Liu_2014}, and Weber and Kirchner \cite{Weber_2016} are, respectively, 1.0\% , 6.2\%, 1.8\%, and 0.5\%.
With the exception of the force field introduced by Batista \textit{et al}. \cite{Batista_2015}, the discrepancies are small.
In all cases, we tried to carefully follow the simulation setup reported by those authors, except for the number of ion pairs.
For instance, Koller and co-workers \cite{Koller_2012} simulated 58 ions pairs, while we considered 250 ion pairs.
Small differences are then expected.
The discrepancy with the results of Batista and co-workers \cite{Batista_2015} stands out from the others.
In this case, however, important computational details are missing in the original paper, such as the method for computing electrostatic interactions, as well as the factors for scaling the 1-4 non-bonded interactions. 
Considering these various factors, we did not focus on reproducing the results of these references, but rather to use the force field parameters reported by the corresponding authors in order to find those that best reproduce the experimental Henry constant of \ce{CO_2}.

@@

{\color{blue}{4) Simulating hybrid systems that include both rigid and non rigid bodies is complex and tricky. The authors should describe in detail how they performed these simulations in terms of barostatting and thermostatting.}}

\textbf{Answer}: On page 5 of the manuscript, we specify the numerical integrators used to carry out the MD simulations.
In fact, the software package we employed for the simulations, LAMMPS, is already prepared for dealing with hybrid systems containing flexible and rigid molecules.
For this purpose, it allows us to define groups of atoms and apply different integrators and constraints to each group.
When this is done, the action of a barostat will scale the positions of all atoms that pertain to flexible molecules, but it will only scale the center-of-mass positions of rigid molecules (or rigid atom groups, in general).

{\color{blue}{5) It would be interesting if at least for the case of CO2 an isotherm was calculated and included in the paper.}}

\textbf{Answer}: We agree with the reviewer.
However, with the methodology we adopted, such computation would require a large number of simulations to be carried out at several pressures and several coupling parameter values for every different number of solute molecules.
Of course, other approaches could be employed for this type of investigation, such as Grand Canonical Monte Carlo simulations, for instance.
While we agree that including isotherms would be of interest, the effort required for doing it would be exceedingly large.
Moreover, we believe that, without obtaining full isotherms, we were able to answer important questions regarding the physical nature of the different \ce{CO_2} solubilities in \ce{[emim][B(CN)_4]} and \ce{[emim][NTF_2]}.
Thus, we ask both the reviewer and the editor to consider publishing our findings without adding these isotherms.

@@

\section{\textbf{Reviewer \#2}}

{\color{blue}{In the present manuscript, Silveira \textit{et al.} performed a systematic comparison of force fields used in Molecular Dynamic (MD) simulations of the solvation of carbon dioxide (\ce{CO_2}) and other compounds in two ionic liquids (ILs). For this purpose, solvation energies, Henry constants, and activity coefficients were calculated based on solid methodologies. A detailed review of the literature with respect to the theoretical background and the force fields is provided. Simplifications in the simulation strategies are analyzed by treating the IL ions as rigid instead of flexible bodies. A structural analysis of the molecular systems and their relationships with the studied properties are investigated thoroughly from a physico-chemical point of view. The authors could demonstrate that the IL \ce{[emim][B(CN)_4]} shows the largest solubility for \ce{CO_2} in comparison to other common ILs. Overall, the scientific findings are very clear, concise, and of high quality, and are presented in a good English. Formalities according to the guidelines of the journal are also fulfilled. In conclusion, the present manuscript is definitely worthy of publication taking into consideration a few minor comments from my side.}}

\textbf{Answer}: We sincerely thank the reviewer for the kind comments and helpful suggestions. They have all been considered and are answered below.

@@

{\color{blue}{1. In the context of the simulation protocol, information on the number simulation runs and the use of the mixing rules for the Lennard-Jones interactions should be given. The authors should also specify how the uncertainties of the various properties were calculated.}}

\textbf{Answer}: It is true that such information was missing and we thank the reviewer for pointing this out.
We have included, on page 6, the total simulation time executed for each system (instead of the number of runs).
Information about the employed mixing rules and uncertainty calculations are now included on pages 5, 9, and 12.
Please note that these page numbers refer to the annotated version of the manuscript.

@@

{\color{blue}{2. The authors were focusing on several static properties for the pure ILs using different models. As reference property for force field validation, the density was selected where deviations between about (1 and 5) \% were found according to Table 1. Did the authors also study dynamic properties such as the self-diffusion coefficient and compare the results with the experiment? If so, a corresponding brief paragraph should be included in the revised manuscript. Relative deviations of 5\% in the liquid density can often come along with deviations of several dozens of percent for dynamic properties.
}}

\textbf{Answer}: Unfortunately, we have not performed any study of dynamic properties.
We have mainly focused on predicting solvation properties, given our goal of investigating differences in the ability of ionic liquids to absorb \ce{CO_2}.
We agree, however, that being able to predict transport properties would be very important for the design of actual separation processes.
It is, therefore, an important subject to be considered in future investigations.

{\color{blue}{3. The results for the simulated solvation free energies of benzene, hexane, and water showed relatively large deviations from experimental data in the order of 10\%, while the results for ethanol match. However, for the infinite-dilution activity coefficient of these solutes in the IL [emim][B(CN)4], hexane deviates significantly stronger than benzene, ethanol, and water. It would be interesting to discuss in some more detail why these different trends can be found by considering, for example, the force fields of the pure solute and the mixing rules for the binary mixtures. For sure, it seems that the authors will continue such investigations in their future studies.}}

\textbf{Answer}: The reviewer is totally right.
In order to analyze the strong deviation observed for hexane, we have included a new comparison of solvation free energies of this and other solutes in \ce{[emim][B(CN)_4]} computed exclusively from experimental data and computed from simulations.
This is shown on page 14, where a new figure (Fig. 7) can be found.
The simulation results predict a favorable solvation of hexane, which contradicts the experimental data.
This is why we emphasize in the conclusions that a future revision of the models used for the solutes is necessary.

@@

{\color{blue}{4. For the specification of the pressure, the symbol “p” instead of “P” should be used. Furthermore, the unit of the pressure should be given in “MPa” instead of “atm."}}

\textbf{Answer}: We accepted the reviewer's suggestion and made the necessary modifications throughout the manuscript.

### 2019-02-19 ionic_liquids.tex
Alpha version for resubmission

Two additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems we investigate here) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equilibrium pressure}
y_s \exp\left( \frac{B_s p}{R T} \right) p = c_s(p) R T \exp \left( \frac{\Delta G_\text{solv}}{R T} \right),
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

@@

In Table \ref{table:henry}, we present results from expanded ensemble simulations of a single \ce{CO_2} molecule dissolved in 250 ion pairs of either \ce{[emim][B(CN)_4]} or \ce{[emim][NTf_2]} under periodic boundary conditions.
The simulations were carried out at $T = 298.0~\mathrm{K}$ and $p = 0.1~\mathrm{MPa}$.
Table~\ref{table:henry} contains simulated molar densities, partial and total solvation free energies, as well as Henry's law constants ($K_s$) computed via Eq.~\eqref{eq:henry_eq}. These should be compared to the experimental data reported by Mahurin \textit{et al}. \cite{Mahurin_2010} ($K_s = 3.941 \pm 0.003 ~\mathrm{MPa}$ for \ce{CO_2} in \ce{[emim][B(CN)_4]}) and by Finotello \textit{et al}. \cite{Finotello_2008} ($K_s = 3.95 \pm 0.01 ~\mathrm{MPa}$ for \ce{CO_2} in \ce{[emim][NTf_2]}).
Uncertainties in $K_s$ were calculated using a standard error propagation formula, given by
\begin{equation}
\delta K_s = \beta K_s \sqrt{\left[ \delta^2 (\Delta G_{\text{vdW}}) + \delta^2(\Delta G_{\text{Coul}}) \right]},
\end{equation}
with the uncertainties in free energies calculated by MBAR \cite{Shirts_2008}.

@@

In order to obtain $\Delta G_\text{sim} (p)$ and $c_s(p)$, we carried out extra simulations at pressure values  of p = (1.722, 1.823) MPa  for \ce{CO_2}+\ce{[emim][B(CN)_4]} and at p = (4.053, 4.458) MPa for \ce{CO_2}+\ce{[emim][NTf_2]}.
In this case, we considered a wider range of pressure, given the departure from ideal gas behavior mentioned above.
As described in Sec.~\ref{sec:partial_pressure}, $\Delta G_\text{sim}$ and $c_s$ were interpolated to intermediate pressures by reweighting using MBAR \cite{Shirts_2008}.
In Fig.~\ref{fig:pressure}, we show for both systems the results of $\Delta G_\text{sim}$, together with the curve of $\Delta G_\text{solv}$ defined by Eq.~\eqref{eq:equilibrium pressure}.
Although not shown here, the changes in density are negligible: 0.05 \% and 0.07 \% for \ce{CO_2}+\ce{[emim][B(CN)_4]} and \ce{CO_2}+\ce{[emim][NTf_2]}, respectively.
In addition, $\Delta G_\text{sim}$ does not vary appreciably with pressure, at least over the range of experimental data considered here. 
This results in small corrections to the equilibrium pressures of Table~\ref{table:solv}: the intersection points are 1.807 MPa and 4.106 MPa, for \ce{CO_2}+\ce{[emim][B(CN)_4]} and \ce{CO_2}+\ce{[emim][NTf_2]}, respectively.
These values did not depart considerably from the specified values, meaning that our results sustain the experimental observations.
Therefore, as the pressure required to reach the same equilibrium molal concentration of \ce{CO_2} in the liquid phase is smaller for \ce{[emim][B(CN)_4]} than for \ce{[emim][NTf_2]}, our results confirm that the former is a better solvent than the latter.
This confirmation was possible by simulating a purely physical absorption, without considering any quantum effect.

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental Henry's law constant of \ce{CO_2}, we extend its use to the calculation of infinite-dilution activity coefficients ($\gamma^\infty$) of other solutes in \ce{[emim][B(CN)_4]}.
Due to the availability of experimental data \cite{Yan_2010, Doma_ska_2011}, the solutes chosen for this study are benzene, hexane, cyclohexane, ethanol, and water. Uncertainties in activity coefficients were calculated using error propagation, such as done for $K_s$.

@@

In Fig.~\ref{fig:deltag_solutes}, we show the simulated solvation free energies versus the experimental results. These were obtained by using $\Delta G^{0}_\text{exp}$ of Table~\ref{table:mu_solutes} and $\gamma^\infty$  from Ref. \cite{Doma_ska_2011} as inputs in Eq.~\eqref{eq:gamma}. Cyclohexane was not considered, given that $\Delta G^{0}_\text{exp}$ is not available.
For benzene, ethanol and water the results from simulations are consistent with the negative experimental values. However, for hexane the simulations predict a favorable solvation, in clear contradiction with the experimental data.
From the individual contributions of $\Delta G_\text{vdW}$ and $\Delta G_\text{Coul}$, we can conclude that the vdW contribution is largely underestimated.
Also, in the case of hexane, one can see in Table \ref{table:gamma} that the available experimental values of $\gamma^\infty$ in \ce{[emim][B(CN)_4]} at very close temperatures ($298.15~\text{K}$ and $303~\text{K}$) differ considerably.
It is, therefore, difficult to assess the reliability of our results in this case.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{delta_g_sim_exp}
	\caption{Simulated free energy $\Delta G_\text{sim}$ versus the experimental result at $T = 298.15~\text{K}$ for the solutes hexane, benzene, ethanol and water in \ce{[emim][B(CN)_4]}.}
	\label{fig:deltag_solutes}
\end{figure}

@@

One of the goals of this paper was to compare various available molecular models of \ce{[emim][B(CN)_4]} in terms of their capabilities to reproduce the experimental Henry's law constant of \ce{CO_2}.
We have found that only the force field developed by Koller \textit{et al}. \cite{Koller_2012} is able to perform such task, with a small relative error of 2.8\%.
A key factor for its success seems to be the reparametrization of the LJ parameters for the boron atom carried out in the original work, which also allows to accurately reproduce the experimental density of the pure substance.

@@

The solvation free energy results suggest that a more favorable solvation of \ce{CO_2} occurs in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}, despite the similarity in the Henry's law constants observed in both cases.
This picture is further confirmed at high \ce{CO_2} concentrations, in which the simulations correctly predicted that a considerably lower pressure is required to dissolve a certain amount of \ce{CO_2} in \ce{[emim][B(CN)_4]}, as compared to \ce{[emim][NTf_2]} in an equal-mass basis.

>>>>

Two additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems we investigate here) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equilibrium pressure}
y_s \exp\left( \frac{B_s p}{R T} \right) p = c_s R T \exp \left( \frac{\Delta G_\text{solv}}{R T} \right),
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

@@

In Table \ref{table:henry}, we present results from expanded ensemble simulations of a single \ce{CO_2} molecule dissolved in 250 ion pairs of either \ce{[emim][B(CN)_4]} or \ce{[emim][NTf_2]} under periodic boundary conditions.
The simulations were carried out at $T = 298.0~\mathrm{K}$ and $p = 0.1~\mathrm{MPa}$.
Table~\ref{table:henry} contains simulated molar densities, partial and total solvation free energies, as well as Henry's law constants ($K_s$) computed via Eq.~\eqref{eq:henry_eq}. These should be compared to the experimental data reported by Mahurin \textit{et al}. \cite{Mahurin_2010} ($K_s = 3.941 \pm 0.003 ~\mathrm{MPa}$ for \ce{CO_2} in \ce{[emim][B(CN)_4]}) and by Finotello \textit{et al}. \cite{Finotello_2008} ($K_s = 3.95 \pm 0.01 ~\mathrm{MPa}$ for \ce{CO_2} in \ce{[emim][NTf_2]}).
Uncertainties in $K_s$ were calculated using a standard error propagation formula, given by
\begin{equation}
\delta K_s = \beta K_s \sqrt{\left[ \delta^2 (\Delta G_{\text{vdW}}) + \delta^2(\Delta G_{\text{Coul}}) \right]},
\end{equation}
with the free-energy uncertainties computed by MBAR \cite{Shirts_2008}.

@@

As one can see in Table~\ref{table:solv}, the estimated equilibrium pressure differs from the specified value for both ionic liquids.
In order to obtain $\Delta G_\text{sim}(p)$ and $c_s(p)$ so that the proper corrections could be made, we carried out additional simulations at some nearby pressures in each case.
Then, new results were obtained at $1.722$ and $1.823~\mathrm{MPa}$ for \ce{CO_2}+\ce{[emim][B(CN)_4]} and at $4.053$ and $4.458~\mathrm{MPa}$ for \ce{CO_2}+\ce{[emim][NTf_2]}.
These ranges were selected after carrying out some short, exploratory simulations.
%In this case, we considered a wider range of pressure, given the departure from ideal gas behavior mentioned above.
As described in Sec.~\ref{sec:partial_pressure}, $\Delta G_\text{sim}$ and $c_s$ were interpolated along the whole pressure ranges by doing reweighting via MBAR \cite{Shirts_2008}.
Each MBAR analysis involved 60 simulation samples (20 coupling parameter states at 3 distinct pressures).
In Fig.~\ref{fig:pressure}, we show the results of $\Delta G_\text{sim}(p)$ for both systems, together with the corresponding curves of $\Delta G_\text{solv}(p)$ defined by Eq.~\eqref{eq:equilibrium pressure}.
The observed variations in density throughout the pressure ranges (results not shown) were negligible: 0.05\% and 0.07\% for \ce{CO_2}+\ce{[emim][B(CN)_4]} and \ce{CO_2}+\ce{[emim][NTf_2]}, respectively.
It is also clear from Fig.~\ref{fig:pressure} that $\Delta G_\text{sim}$ does not vary appreciably either.
As a result, the corrections required for the equilibrium pressures shown in Table~\ref{table:solv} are small.
The intersection points in the mentioned curves occur at 1.807 MPa for \ce{CO_2}+\ce{[emim][B(CN)_4]} and at 4.106 MPa \ce{CO_2}+\ce{[emim][NTf_2]}, meaning that our results corroborate the experimental observations.
Therefore, as the pressure required to reach the same equilibrium molal concentration of \ce{CO_2} in the liquid phase is smaller for \ce{[emim][B(CN)_4]} than for \ce{[emim][NTf_2]}, our results confirm that the former is a better solvent than the latter.
This confirmation was possible by simulating a purely physical absorption, without considering any quantum effect.

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental Henry's law constant of \ce{CO_2}, we extend its use to the calculation of infinite-dilution activity coefficients ($\gamma^\infty$) of other solutes in \ce{[emim][B(CN)_4]}.
Due to the availability of experimental data \cite{Yan_2010, Doma_ska_2011}, the solutes chosen for this study are benzene, hexane, cyclohexane, ethanol, and water. Uncertainties in activity coefficients were evaluated by error propagation, such as done for Henry's law constants.

@@

In Fig.~\ref{fig:deltag_solutes}, we compare solvation free energies obtained by simulation with those computed from experimental results.
These were obtained by using $\Delta G^{0}_\text{exp}$ of Table~\ref{table:mu_solutes} and $\gamma^\infty$ from Ref. ~\citenum{Doma_ska_2011} as inputs to Eq.~\eqref{eq:gamma}.
Because $\Delta G^{0}_\text{exp}$ is not available for cyclohexane, such substance was not considered in the comparison.
For benzene, ethanol, and water, the results from simulations are consistent with the negative experimental values. For hexane, however, the simulation prediction of a favorable solvation is in clear contradiction with the experimental data.
From the individual contributions of $\Delta G_\text{vdW}$ and $\Delta G_\text{Coul}$, we can conclude that the van der Waals contribution is largely underestimated.
Also, in the case of hexane, one can see in Table~\ref{table:gamma} that the available experimental values of $\gamma^\infty$ in \ce{[emim][B(CN)_4]} at very close temperatures ($298.15~\text{K}$ and $303~\text{K}$) differ considerably.
It is, therefore, difficult to assess the reliability of our results in this case.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{delta_g_sim_exp}
	\caption{Solvation free energies at infinite dilution obtained from simulations ($\Delta G_\text{sim}$) in comparison to values computed from experimental data ($\Delta G_\text{exp}$) for the solutes hexane, benzene, ethanol, and water in \ce{[emim][B(CN)_4]}  at $T = 298.15~\text{K}$.}
	\label{fig:deltag_solutes}
\end{figure}

@@

One of the goals of this paper was to compare various available molecular models of \ce{[emim][B(CN)_4]} in terms of their capabilities to reproduce the experimental Henry's law constant of \ce{CO_2}.
We have found that only the force field developed by Koller \textit{et al}. \cite{Koller_2012} is able to perform such task, with a small relative error of 2.8\%.
A key factor for its success seems to be the reparametrization of the LJ parameters for the boron atom carried out in the original work, which also allows us to accurately reproduce the experimental density of the pure substance.

@@

The solvation free energy results at infinite dilution condition suggest that a more favorable solvation of \ce{CO_2} occurs in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}, despite the similarity in the Henry's law constants observed in both cases.
This picture is further confirmed at high \ce{CO_2} concentrations, in which the simulations correctly predicted that a considerably lower pressure is required to dissolve a certain amount of \ce{CO_2} in \ce{[emim][B(CN)_4]}, as compared to \ce{[emim][NTf_2]} in an equal-mass basis.
